---
title: "Hoa_Nguyen_casetudy_cyclistic"
author: "Hoa Nguyen"
date: "2023-04-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This is my analysis of the case study of the **Google Data Analytics** certificate program. I'll be performing my process of data cleaning, analyzing, and visualizing data, then summarizing the data and delivering insights to solve business questions.

# Scenario

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director
of marketing believes the company’s future success depends on maximizing the number of annual memberships. Therefore,
your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights,
your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives
must approve your recommendations, so they must be backed up with compelling data insights and professional data
visualizations.

# Characters and teams

* **Cyclistic:** A bike-share program that features more than 5,800 bicycles and 600 docking stations. Cyclistic sets itself
apart by also offering reclining bikes, hand tricycles, and cargo bikes, making bike-share more inclusive to people with
disabilities and riders who can’t use a standard two-wheeled bike. The majority of riders opt for traditional bikes; about
8% of riders use the assistive options. Cyclistic users are more likely to ride for leisure, but about 30% use them to
commute to work each day.
* **Lily Moreno:** The director of marketing and your manager. Moreno is responsible for the development of campaigns
and initiatives to promote the bike-share program. These may include email, social media, and other channels.
* **Cyclistic marketing analytics team:** A team of data analysts who are responsible for collecting, analyzing, and
reporting data that helps guide Cyclistic marketing strategy. You joined this team six months ago and have been busy
learning about Cyclistic’s mission and business goals — as well as how you, as a junior data analyst, can help Cyclistic
achieve them.
* **Cyclistic executive team:** The notoriously detail-oriented executive team will decide whether to approve the
recommended marketing program.

I'll be following the six phases of data analysis: 
* Ask
* Prepare
* Process
* Analyze
* Share
* Act

# Ask
*In this phase, we define the problem and understand the stakeholder expectations*
The problem of this case study **1.How do annual members and casual riders use Cyclistic bikes differently?
2. Why would casual riders buy Cyclistic annual memberships?
3. How can Cyclistic use digital media to influence casual riders to become members?**
the stakeholder expectations: *Design marketing strategies aimed at converting casual riders into annual members*

# Prepare
*In this phase we will collect and store data then use for upcomming analysis process.*

*Identify which kinds of data are most useful for solving a particular problem.*

The data can be downloaded at \textcolor{blue}{}[divvy_trip](https://divvy-tripdata.s3.amazonaws.com/index.html)

I will use the most recent year of data based on my current time.
The data time frame is from **1/2022 to 12/2022**. The data is separated by each month in each file.

This data is suitable for solving this business problem because it contains insights into riders' patterns of behavior.
## Load data 

First, we will load the package needed for the data analysis process.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(skimr)
library(hydroTSM)
library(dplyr)
```

```{r}
M1<-read.csv("C:/Users/ADMIN/Videos/Captures/202201-divvy-tripdata.csv")
M2<-read.csv("C:/Users/ADMIN/Videos/Captures/202202-divvy-tripdata.csv")
M3<-read.csv("C:/Users/ADMIN/Videos/Captures/202203-divvy-tripdata.csv")
M4<-read.csv("C:/Users/ADMIN/Videos/Captures/202204-divvy-tripdata.csv")
M5<-read.csv("C:/Users/ADMIN/Videos/Captures/202205-divvy-tripdata.csv")
M6<-read.csv("C:/Users/ADMIN/Videos/Captures/202206-divvy-tripdata.csv")
M7<-read.csv("C:/Users/ADMIN/Videos/Captures/202207-divvy-tripdata.csv")
M8<-read.csv("C:/Users/ADMIN/Videos/Captures/202208-divvy-tripdata.csv")
M9<-read.csv("C:/Users/ADMIN/Videos/Captures/202209-divvy-publictripdata.csv")
M10<-read.csv("C:/Users/ADMIN/Videos/Captures/202210-divvy-tripdata.csv")
M11<-read.csv("C:/Users/ADMIN/Videos/Captures/202211-divvy-tripdata.csv")
M12<-read.csv("C:/Users/ADMIN/Videos/Captures/202212-divvy-tripdata.csv")
```

# Compare column names each of the files
# We match column names perfectly before we can use a command to join them into single file
```{r}
colnames(M1)
colnames(M2)
colnames(M3)
colnames(M4)
colnames(M5)
colnames(M6)
colnames(M7)
colnames(M8)
colnames(M9)
colnames(M10)
colnames(M11)
colnames(M12)
```
**\textcolor{yellow}{All the column names of files matched each other}
# We combine all the files into a single dataframe
```{r}
data_combined <- bind_rows(M1, M2, M3, M4, M5, M6, M7, M8, M9, M10, M11, M12)
colnames(data_combined)
```
We remove the useless variables and take the rest. Convert irrelevant data type(character into Date data type)

```{r}
data_combined <- select(data_combined, -c(ride_id, start_station_id, end_station_id, start_lat, start_lng, end_lat, end_lng))
data_combined <- mutate(data_combined, started_at = as.POSIXct(started_at, format = "%Y-%m-%d %H:%M:%S"), ended_at = as.POSIXct(ended_at, format = "%Y-%m-%d %H:%M:%S"))
colnames(data_combined)
```
Dimensions of the dataframe

```{r}
dim(data_combined)
```
# Process

*We will find and eliminate any error and inaccuracy in the data.*


Add new columns(date, day, month, year, day of the week, season, ride_length)

```{r}
data_Process <- data_combined
data_Process$date <- as.Date(data_Process$started_at)
data_Process$day <- format(as.Date(data_Process$date), '%d')
data_Process$month <- format(as.Date(data_Process$date), '%m')
data_Process$year <- format(as.Date(data_Process$date), '%Y')
data_Process$day_of_week <- format(as.Date(data_Process$date), '%A')
data_Process$season <- time2season(as.Date(data_Process$date), out.fmt = "seasons")
```
```{r}
data_Process$ride_length <- difftime(data_Process$ended_at,data_Process$started_at, units = 'mins')
data_Process$ride_length <- round(data_Process$ride_length, digits = 1)
```
Change ride_length data type to numeric.

```{r}
data_Process$ride_length <- as.numeric(data_Process$ride_length)
```
We will be cleaning data next. We remove rows with ride_length below or equal to 0.


```{r}
data_Process <- data_Process[!(data_Process$ride_length <= 0),]
```
```{r}
skim_without_charts(data_combined)
```


```{r}
summary(data_Process)
```

```{r}
nrow(filter(data_Process, data_Process$start_station_name == ''))
```
```{r}
summary(data_Process$ride_length)
```
Calculate mean by member casual

```{r}
data_Process %>%
  group_by(member_casual) %>%
  summarise(mean = mean(ride_length))
```
```{r}
nrow(filter(data_Process, data_Process$member_casual == 'casual'))/nrow(data_Process)
nrow(filter(data_Process, data_Process$member_casual == 'member'))/nrow(data_Process)
```

write.csv(data_Process, file = "dong.csv")
write.csv(data_Process, file = "hoa.csv")
# Analyze
*For this phase we will be using tools to transform and organize information so we can draw useful information/conclusion.*

We can use R to visualize data and transform it in any way we like. But I would like to use Tableau as practice because it is also part of the course and I want to make use of every tool I was taught.(
For the Tableau Dashboard you can visit [Here] (https://public.tableau.com/app/profile/nguy.n.hoa8602/viz/Cyclistic_case_study_16822512065760/Tableaudebord1)

After visualization, we can infer some information:

*  Members take **59%** of total rides. 
* The most **popular bike types** are classic and electric. For casual riders, they prefer docked bikes more than member riders.
* Riders **ride the most** on **Saturday**. Member riders ride equally throughout the weekdays with slight increase in mid weekdays. Casual riders ride equally on weekdays and more on weekends.
* **Summer** is the **most busy** season for both riders type and **Winter** is the **least busy** season for both riders type.

# Share

*We will interpret result and share with others to help stakeholders making data-driven-decision*

We will demonstrate and make a presentation to the stakeholders to deliver our findings in order to help them make a decision on the problem.

# Act

* To turn casual riders into members, we can make discounts on the time when riders ride the most like Summer or on the weekend.
* Show them the perks of becoming a member, and customize the discount and membership program for their specific riding habits.